Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-03-11 Manuscript/manuscript.Rmd
manuscript edits

>>>>

## MODELS WITH DISTANCE-TYPE PREDICTORS OUTPERFORM MODELS WITH OTHER PREDICTOR-TYPES

After suitably reducing the feature space, we performed individual random forest classification models specific to cell line, data resolution, resampling technique, and predictor type. We assessed each models' ability to correctly identify the genomic bins that overlap with flanked TAD boundaries using F1-Score, MCC, and balanced accuracy. For 10 kb data resoltion, it was found that models with distance-type predictors performed better according to all metrics, across each cell line, regardless of resampling technique used (Figure 4). This result was consistent across each of the other data resolutions as well (Supplementary Figure 4). Notably, we saw marked improvements in performance for models implemented on resampled datasets when compared to no resampling. Of the, resampling techniques considered, it was found that SMOTE resampled datasets, at the 10 kb resoltion, yielded the highest model performances across each metric. This result was, for the most part, consistent for the other data resolutions. Performances of randomly under-sampled datasets were found to be comparable to, if not better than (at some isolated resolutions; see supplementary), SMOTE. The steap drop off in performance for randomly over-sampled and no resampled datasets was most evident when assessing F1-scores and balanced accuracies. Measures of MCC were found to increase more steadily toward models implemented on SMOTE resampled datasets. These results imply that a best performing model, one that is able to best discriminate between genomic bins that are near versus far away from TAD boundaries, is one that incorporates both random under- and over-sampling, as SMOTE is known to do.
